Alex's Notes and Text Fragments for the JSS Manuscript

### 2024-01-05 paper/article.Rmd
add knndm section and table row

```{r setup-2, echo=FALSE, purl=TRUE}
knitr::opts_chunk$set(
  cache = FALSE,
  cache.lazy = TRUE)
```

@@

\begin{table}[ht]
  \centering
  \footnotesize
  \begingroup
    \begin{tabular}{lllll}
      \\
      Type                            & Sub-type & Name                        & \proglang{R} package   & References \\
      \toprule
      \multirow{4}{*}{\specialcell{Spatial \\ leave-one-\\out}}        & single point, with buffer &  \texttt{"spcv\_buffer"} & \pkg{blockCV} (2) & \specialcell{\cite{ploton2020} \\ \cite{diesing2020}} \\
      \cmidrule{2-5}
       & disc, with buffer  & \texttt{"spcv\_disc"}  & \pkg{sperrorest} (3) & \specialcell{\cite{karasiak2021} \\ \cite{moller2021} \\ \cite{endicott2017}} \\
      \midrule
      \multirow{4}{*}{\specialcell{Leave-one-\\block-out \\ CV}} &  clustering of coordinates & \texttt{"spcv\_coords"}* & \pkg{sperrorest} (6)      & \specialcell{\cite{morera2021} \\ \cite{geiss2017} \\ \cite{wu2020}} \\
            \cmidrule{2-5}
                                      &  geometric: rectangular & \texttt{"spcv\_tiles"} & \pkg{sperrorest}    & \specialcell{\cite{bebber2017} \\ \cite{zurell2020} \\ \cite{brenning2015}}  \\
                                                  \cmidrule{2-5}
                                      & custom  &  \texttt{"custom\_cv"}           & \pkg{mlr3} (0) & - \\
      \midrule
      \multirow{4}{*}{\specialcell{CV at the \\ block level}} &  geometric: rectangular & \texttt{"spcv\_block"}  & \pkg{blockCV} (28)      & \specialcell{\cite{jensen2021} \\ \cite{escobar2021} \\ \cite{stewart2021}} \\
  \cmidrule{2-5}
                                      &  custom & \texttt{"cv"} with grouping  & \pkg{mlr3} (0)  &  -  \\
  \cmidrule{2-5}
                                      &  clustering in feature space & \texttt{"spcv\_env"}* & \pkg{blockCV} (1)   &  \cite{morera2021}  \\
      \midrule
      \multirow{2}{*}{\specialcell{Spatiotemp. \\ CV}} &  custom & \texttt{"sptcv\_cstf"}* & \pkg{CAST} (6)      & \specialcell{\cite{gao2019} \\ \cite{reitz2021} \\ \cite{egli2020}} \\
                                      \bottomrule
    \end{tabular}
  \endgroup
  \caption[b]{Available spatiotemporal resampling methods in the \pkg{mlr3} ecosystem.
    The "Name" column shows the \pkg{mlr3} method name as found in the \texttt{mlr3::mlr\_resamplings} dictionary.
    The count in brackets after the package name represents the number of studies that were found having used this resampling technique until May 2021.
    For each method, up to three randomly selected references were added to the table.
    Methods suffixed with a * in the "Name" column have been reimplemented in \pkg{mlr3spatiotempcv} for efficiency or to reduce upstream dependencies.}
    \label{tab:sptcv-methods}
\end{table}

@@

```{r buffer-eval, fig.cap='Visualization of the spatial buffering method from package \\pkg{blockCV} (method "spcv\\_buffer" in \\pkg{mlr3spatiotempcv}). The buffer distance is 1000 m.', out.width="60%", fig.pos="ht",echo=FALSE}
library("mlr3")
library("mlr3spatiotempcv")
task = tsk("ecuador")
rsmp_buffer = rsmp("spcv_buffer", theRange = 1000)
rsmp_buffer

autoplot(rsmp_buffer,
  size = 0.8, task = task, fold_id = 1, show_omitted = TRUE) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.02)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.02)) +
  ggplot2::theme(text = ggplot2::element_text(size = 16),
    plot.title = ggtext::element_textbox(size = 13,
      r = ggplot2::unit(7, "pt"),
      height = ggplot2::unit(0.46, "inch"), linewidth = 0.8)) +
  ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size = 1.5)))
```

@@

autoplot(rsmp_disc,
  size = 0.8, task = task, fold_id = 1, show_omitted = TRUE) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.02)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.02)) +
  ggplot2::theme(text = ggplot2::element_text(size = 16),
    plot.title = ggtext::element_textbox(size = 13,
      r = ggplot2::unit(7, "pt"),
      height = ggplot2::unit(0.46, "inch"), linewidth = 0.8)) +
  ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size = 1.5)))
```

@@

autoplot(rsmp_coords,
  size = 0.8, fold_id = 1, task = task) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.02)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.02)) +
  ggplot2::theme(text = ggplot2::element_text(size = 16),
    plot.title = ggtext::element_textbox(size = 13,
      r = ggplot2::unit(7, "pt"),
      height = ggplot2::unit(0.46, "inch"), linewidth = 0.8)) +
  ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size = 1.5)))
```

### Clustering-based: using k-fold nearest-neighbour distance matching --- \code{"spcv_knndm"} {short-title='Clustering-based: using kNNDM --- "spcv_knndm"' }

The k-fold nearest-neighboar distance matching method after @linnenbrinck2023 is derived from the NNDM method after @mila.
By using geographical nearest neighbor distances between test and training points, the method aims to create realistic predictive situations within the CV folds that closely mimic the ones of either existing training and test points, if know, or a provided model domain in which the model is to be applied.
This is done by minimizing the values ofthe Empirical Cumulative Distribution Function (ECDF) for the respective scenarios outlined above [@linnenbrinck2023].

```{r knndm-non-eval, eval=FALSE}
rsmp_knndm = rsmp("spcv_knndm", folds = 5)

autoplot(rsmp_knndm, size = 0.8, fold_id = 1, task = task)
```

```{r knndm-eval, fig.cap='Leave-one-block-out CV based on $k$-means clustering of the coordinates as implemented in package \\pkg{sperrorest} (method `"spcv\\_coords"` in \\pkg{mlr3spatiotempcv}).', out.width="55%", fig.pos="ht", echo=FALSE}
rsmp_coords = rsmp("spcv_coords", folds = 5)

autoplot(rsmp_coords,
  size = 0.8, fold_id = 1, task = task) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.02)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.02)) +
  ggplot2::theme(text = ggplot2::element_text(size = 16),
    plot.title = ggtext::element_textbox(size = 13,
      r = ggplot2::unit(7, "pt"),
      height = ggplot2::unit(0.46, "inch"), linewidth = 0.8)) +
  ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size = 1.5)))
```

FIXME:

@@

autoplot(rsmp_tiles,
  size = 0.8, fold_id = 1, task = task) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.02)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.02)) +
  ggplot2::theme(text = ggplot2::element_text(size = 16),
    plot.title = ggtext::element_textbox(size = 13,
      r = ggplot2::unit(7, "pt"),
      height = ggplot2::unit(0.46, "inch"), linewidth = 0.8)) +
  ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size = 1.5)))
```

@@

autoplot(rsmp_custom, size = 0.8, task = task, fold_id = 1) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.02)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.02)) +
  ggplot2::theme(text = ggplot2::element_text(size = 16),
    plot.title = ggtext::element_textbox(size = 13,
      r = ggplot2::unit(7, "pt"),
      height = ggplot2::unit(0.46, "inch"), linewidth = 0.8)) +
  ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size = 1.5)))
```

@@

autoplot(rsmp_block_random, size = 0.8, fold_id = 1, task = task,
  show_blocks = TRUE, show_labels = TRUE, label_size = 4)
```

@@

autoplot(rsmp_block_random,
  size = 0.8, fold_id = 1, task = task,
  show_blocks = TRUE, show_labels = TRUE, label_size = 4) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.02)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.02)) +
  ggplot2::theme(text = ggplot2::element_text(size = 16),
    plot.title = ggtext::element_textbox(size = 13,
      r = ggplot2::unit(7, "pt"),
      height = ggplot2::unit(0.46, "inch"), linewidth = 0.8)) +
  ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size = 1.5)))
```

@@

autoplot(rsmp_block_systematic, size = 0.8, fold_id = 1, task = task,
  show_blocks = TRUE, show_labels = TRUE, label_size = 4)
```

@@

autoplot(rsmp_block_systematic,
  size = 0.8, fold_id = 1, task = task,
  show_blocks = TRUE, show_labels = TRUE, label_size = 4) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.02)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.02)) +
  ggplot2::theme(text = ggplot2::element_text(size = 16),
    plot.title = ggtext::element_textbox(size = 13,
      r = ggplot2::unit(7, "pt"),
      height = ggplot2::unit(0.46, "inch"), linewidth = 0.8)) +
  ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size = 1.5)))
```

@@

```{r cv-eval-2, fig.cap='Cross-Validation at the block level including predefined groups from package \\pkg{mlr3} (method `"cv"`). A factor variable is used to define the grouping. Each class is either assigned to the test or training set.', out.width="60%", fig.pos="ht", echo=FALSE}
autoplot(rsmp_cv_group, size = 0.8, task = task_cv, fold_id = 1) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.02)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.02)) +
  ggplot2::theme(text = ggplot2::element_text(size = 16),
    plot.title = ggtext::element_textbox(size = 13,
      r = ggplot2::unit(7, "pt"),
      height = ggplot2::unit(0.46, "inch"), linewidth = 0.8)) +
  ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size = 1.5)))
```

@@

rsmp_env_multi = rsmp("spcv_env", features = c("distdeforest", "slope"),
  folds = 5)

plot_env_single = autoplot(rsmp_env, size = 0.3,
  fold_id = 1, task = task) +

  plot_env_multi = autoplot(rsmp_env_multi, size = 0.3,
  fold_id = 1, task = task)

@@

```{r env-1-eval, fig.cap='Environmental leave-one-block-out CV from package \\pkg{blockCV} using one (left, `"distdeforest"`) and two (right, `"distdeforest"` and `"slope"`) predictors to define blocks in the feature space. Due to feature space clustering observations are not (necessarily) grouped in the spatial domain.', out.width="100%", fig.pos="ht", echo=FALSE}
rsmp_env = rsmp("spcv_env",
  features = "distdeforest", folds = 5)

rsmp_env_multi = rsmp("spcv_env",
  features = c("distdeforest", "slope"), folds = 5)

plot_env_single = autoplot(rsmp_env,
  size = 0.3, fold_id = 1, task = task) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.02)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.02)) +
  ggplot2::theme(plot.title = ggtext::element_textbox(size = 8),
    axis.text = ggplot2::element_text(size = 7),
    legend.title = ggtext::element_textbox(size = 8),
    legend.text = ggtext::element_textbox(size = 6),
    legend.key.size = ggplot2::unit(0.4, "cm"))

plot_env_multi = autoplot(rsmp_env_multi,
  size = 0.3, fold_id = 1, task = task) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.02)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.02)) +
  ggplot2::theme(plot.title = ggtext::element_textbox(size = 8),
    axis.text = ggplot2::element_text(size = 7),
    legend.title = ggtext::element_textbox(size = 8),
    legend.text = ggtext::element_textbox(size = 6),
    legend.key.size = ggplot2::unit(0.4, "cm"))

@@

@@

```{r, results='hide'}
data = cookfarm_mlr3
set.seed(42)
data$Date = sample(rep(c(
  "2020-01-01", "2020-02-01", "2020-03-01", "2020-04-01",
  "2020-05-01"), times = 1, each = 35768))
task_spt = as_task_regr_st(data,
  id = "cookfarm", target = "PHIHOX",
  coordinate_names = c("x", "y"), coords_as_features = FALSE,
  crs = 26911)
task_spt$set_col_roles("Date", roles = "time")

@@

p_lto_print = plotly::layout(p_lto,
  scene = list(camera = list(eye = list(z = 0.58))),
  showlegend = FALSE, title = "",
  margin = list(l = 0, b = 0, r = 0, t = 0))

plotly::save_image(p_lto_print, "pdf/lto.pdf",
  scale = 2, width = 1000, height = 800)
```

@@

p_llo = autoplot(rsmp_cstf_loc,
  fold_id = 5, task = task_spt,
  point_size = 6, axis_label_fontsize = 15,
  plot3D = TRUE, plot_time_var = "Date",
  sample_fold_n = 3000L)

p_llo_print =
  plotly::layout(p_llo,
    scene = list(camera = list(eye = list(z = 2.5, x = -0.1, y = -0.1))),
    showlegend = FALSE, title = "", polar = TRUE,
    margin = list(l = 0, b = 0, r = 0, t = 0))

plotly::save_image(p_llo_print, "pdf/llo.pdf",
  scale = 2, width = 1000, height = 800)
```

@@

p_lto = autoplot(rsmp_cstf_time_loc, point_size = 6,
  axis_label_fontsize = 15,
  fold_id = 4, task = task_spt, plot3D = TRUE,
  show_omitted = TRUE, sample_fold_n = 3000L)

p_lto_print = plotly::layout(p_lto,
  scene = list(camera = list(eye = list(z = 0.58))),
  showlegend = FALSE, title = "",
  margin = list(l = 0, b = 0, r = 0, t = 0))

plotly::save_image(p_lto_print, "pdf/llto.pdf",
  scale = 2, width = 1000, height = 800)
```

@@

```{r create-task}
data("ecuador", package = "mlr3spatiotempcv")
task = as_task_classif_st(ecuador, target = "slides", positive = "TRUE",
  coordinate_names = c("x", "y"), coords_as_features = FALSE,
  crs = "EPSG:32717")
```

@@

```{r vis-spcv-eval, fig.cap="Spatial leave-one-block-out partitioning using coordinate-based clustering to create roughly equally sized polygonal blocks. Due to space limitations only the first two folds of the first repetition are shown.", fig.pos="ht", out.width="100%", echo=FALSE}
autoplot(rsmp_sp, task, fold_id = 1:2, size = 0.3) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.02)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.02)) *
  ggplot2::theme(plot.title = ggtext::element_textbox(size = 8),
    axis.text = ggplot2::element_text(size = 7),
    legend.title = ggtext::element_textbox(size = 8),
    legend.text = ggtext::element_textbox(size = 7),
    legend.key.size = ggplot2::unit(0.4, "cm"))
```

@@

For example, buffering, or the use of a spatial or temporal separation distance between training and test sets, is currently only implemented for some methods (`"spcv_buffer"`, `"spcv_disc"`, and `"sptcv_cstf"` with both `space_var` and `time_var`).
Its use should, however, be limited to use cases involving a prediction distance, as a buffer zone reduces the size of the training sample and introduces the risk of geographically biased training data.

>>>>

```{r setup-2, echo=FALSE, purl=TRUE}
knitr::opts_chunk$set(
  cache = FALSE,
  cache.lazy = TRUE
)
```

@@

\begin{table}[ht]
  \centering
  \footnotesize
  \begingroup
    \begin{tabular}{lllll}
      \\
      Type                            & Sub-type & Name                        & \proglang{R} package   & References \\
      \toprule
      \multirow{4}{*}{\specialcell{Spatial \\ leave-one-\\out}}        & single point, with buffer &  \texttt{"spcv\_buffer"} & \pkg{blockCV} (2) & \specialcell{\cite{ploton2020} \\ \cite{diesing2020}} \\
      \cmidrule{2-5}
       & disc, with buffer  & \texttt{"spcv\_disc"}  & \pkg{sperrorest} (3) & \specialcell{\cite{karasiak2021} \\ \cite{moller2021} \\ \cite{endicott2017}} \\
      \midrule
      \multirow{4}{*}{\specialcell{Leave-one-\\block-out \\ CV}} &  clustering of coordinates & \texttt{"spcv\_coords"}* & \pkg{sperrorest} (6)      & \specialcell{\cite{morera2021} \\ \cite{geiss2017} \\ \cite{wu2020}} \\
            \cmidrule{2-5}
                                      &  geometric: rectangular & \texttt{"spcv\_tiles"} & \pkg{sperrorest}    & \specialcell{\cite{bebber2017} \\ \cite{zurell2020} \\ \cite{brenning2015}}  \\
                                                  \cmidrule{2-5}
                                      & custom  &  \texttt{"custom\_cv"}           & \pkg{mlr3} (0) & - \\
      \midrule
      \multirow{5}{*}{\specialcell{CV at the \\ block level}} &  geometric: rectangular & \texttt{"spcv\_block"}  & \pkg{blockCV} (28)      & \specialcell{\cite{jensen2021} \\ \cite{escobar2021} \\ \cite{stewart2021}} \\
  \cmidrule{2-5}
                                      &  custom & \texttt{"cv"} with grouping  & \pkg{mlr3} (0)  &  -  \\
  \cmidrule{2-5}
                                      &  clustering in feature space & \texttt{"spcv\_env"}* & \pkg{blockCV} (1)   &  \cite{morera2021}  \\
  \cmidrule{2-5}
                                      &  clustering of coordinates + aggregation & \texttt{"spcv\_knndm"}* & \pkg{CAST} (1)   &  \cite{ludwig2023}  \\
      \midrule
      \multirow{2}{*}{\specialcell{Spatiotemp. \\ CV}} &  custom & \texttt{"sptcv\_cstf"}* & \pkg{CAST} (6)      & \specialcell{\cite{gao2019} \\ \cite{reitz2021} \\ \cite{egli2020}} \\
                                      \bottomrule
    \end{tabular}
  \endgroup
  \caption[b]{Available spatiotemporal resampling methods in the \pkg{mlr3} ecosystem.
    The "Name" column shows the \pkg{mlr3} method name as found in the \texttt{mlr3::mlr\_resamplings} dictionary.
    The count in brackets after the package name represents the number of studies that were found having used this resampling technique until May 2021.
    For each method, up to three randomly selected references were added to the table.
    Methods suffixed with a * in the "Name" column have been reimplemented in \pkg{mlr3spatiotempcv} for efficiency or to reduce upstream dependencies.}
    \label{tab:sptcv-methods}
\end{table}

@@

```{r buffer-eval, fig.cap='Visualization of the spatial buffering method from package \\pkg{blockCV} (method "spcv\\_buffer" in \\pkg{mlr3spatiotempcv}). The buffer distance is 1000 m.', out.width="60%", fig.pos="ht",echo=FALSE}
set.seed(42)
library("mlr3")
library("mlr3spatiotempcv")
task = tsk("ecuador")
rsmp_buffer = rsmp("spcv_buffer", theRange = 1000)
rsmp_buffer

autoplot(rsmp_buffer,
  size = 0.8, task = task, fold_id = 1, show_omitted = TRUE
) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.02)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.02)) +
  ggplot2::theme(
    text = ggplot2::element_text(size = 16),
    plot.title = ggtext::element_textbox(
      size = 13,
      r = ggplot2::unit(7, "pt"),
      height = ggplot2::unit(0.46, "inch"), linewidth = 0.8
    )
  ) +
  ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size = 1.5)))
```

@@

autoplot(rsmp_disc,
  size = 0.8, task = task, fold_id = 1, show_omitted = TRUE
) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.02)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.02)) +
  ggplot2::theme(
    text = ggplot2::element_text(size = 16),
    plot.title = ggtext::element_textbox(
      size = 13,
      r = ggplot2::unit(7, "pt"),
      height = ggplot2::unit(0.46, "inch"), linewidth = 0.8
    )
  ) +
  ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size = 1.5)))
```

@@

autoplot(rsmp_coords,
  size = 0.8, fold_id = 1, task = task
) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.02)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.02)) +
  ggplot2::theme(
    text = ggplot2::element_text(size = 16),
    plot.title = ggtext::element_textbox(
      size = 13,
      r = ggplot2::unit(7, "pt"),
      height = ggplot2::unit(0.46, "inch"), linewidth = 0.8
    )
  ) +
  ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size = 1.5)))
```

@@

autoplot(rsmp_tiles,
  size = 0.8, fold_id = 1, task = task
) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.02)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.02)) +
  ggplot2::theme(
    text = ggplot2::element_text(size = 16),
    plot.title = ggtext::element_textbox(
      size = 13,
      r = ggplot2::unit(7, "pt"),
      height = ggplot2::unit(0.46, "inch"), linewidth = 0.8
    )
  ) +
  ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size = 1.5)))
```

@@

autoplot(rsmp_custom, size = 0.8, task = task, fold_id = 1) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.02)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.02)) +
  ggplot2::theme(
    text = ggplot2::element_text(size = 16),
    plot.title = ggtext::element_textbox(
      size = 13,
      r = ggplot2::unit(7, "pt"),
      height = ggplot2::unit(0.46, "inch"), linewidth = 0.8
    )
  ) +
  ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size = 1.5)))
```

@@

autoplot(rsmp_block_random,
  size = 0.8, fold_id = 1, task = task,
  show_blocks = TRUE, show_labels = TRUE, label_size = 4
)
```

@@

autoplot(rsmp_block_random,
  size = 0.8, fold_id = 1, task = task,
  show_blocks = TRUE, show_labels = TRUE, label_size = 4
) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.02)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.02)) +
  ggplot2::theme(
    text = ggplot2::element_text(size = 16),
    plot.title = ggtext::element_textbox(
      size = 13,
      r = ggplot2::unit(7, "pt"),
      height = ggplot2::unit(0.46, "inch"), linewidth = 0.8
    )
  ) +
  ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size = 1.5)))
```

@@

autoplot(rsmp_block_systematic,
  size = 0.8, fold_id = 1, task = task,
  show_blocks = TRUE, show_labels = TRUE, label_size = 4
)
```

@@

autoplot(rsmp_block_systematic,
  size = 0.8, fold_id = 1, task = task,
  show_blocks = TRUE, show_labels = TRUE, label_size = 4
) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.02)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.02)) +
  ggplot2::theme(
    text = ggplot2::element_text(size = 15),
    plot.title = ggtext::element_textbox(
      size = 12,
      r = ggplot2::unit(7, "pt"),
      height = ggplot2::unit(0.46, "inch"), linewidth = 0.8
    )
  ) +
  ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size = 1.5)))
```

@@

```{r cv-eval-2, fig.cap='Cross-Validation at the block level including predefined groups from package \\pkg{mlr3} (method `"cv"`). A factor variable is used to define the grouping. Each class is either assigned to the test or training set.', out.width="60%", fig.pos="ht", echo=FALSE}
autoplot(rsmp_cv_group, size = 0.8, task = task_cv, fold_id = 1) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.02)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.02)) +
  ggplot2::theme(
    text = ggplot2::element_text(size = 16),
    plot.title = ggtext::element_textbox(
      size = 13,
      r = ggplot2::unit(7, "pt"),
      height = ggplot2::unit(0.46, "inch"), linewidth = 0.8
    )
  ) +
  ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size = 1.5)))
```

@@

rsmp_env_multi = rsmp("spcv_env",
  features = c("distdeforest", "slope"),
  folds = 5
)

plot_env_single = autoplot(rsmp_env,
  size = 0.3,
  fold_id = 1, task = task
) +

  plot_env_multi = autoplot(rsmp_env_multi,
  size = 0.3,
  fold_id = 1, task = task
)

@@

```{r env-1-eval, fig.cap='Environmental leave-one-block-out CV from package \\pkg{blockCV} using one (left, `"distdeforest"`) and two (right, `"distdeforest"` and `"slope"`) predictors to define blocks in the feature space. Due to feature space clustering observations are not (necessarily) grouped in the spatial domain.', out.width="100%", fig.pos="ht", echo=FALSE}
rsmp_env = rsmp("spcv_env",
  features = "distdeforest", folds = 5
)

rsmp_env_multi = rsmp("spcv_env",
  features = c("distdeforest", "slope"), folds = 5
)

plot_env_single = autoplot(rsmp_env,
  size = 0.3, fold_id = 1, task = task
) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.02)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.02)) +
  ggplot2::theme(
    plot.title = ggtext::element_textbox(
      size = 8,
      r = ggplot2::unit(4, "pt"),
      height = ggplot2::unit(0.30, "inch"), linewidth = 0.5
    ),
    axis.text = ggplot2::element_text(size = 8.5),
    legend.title = ggtext::element_textbox(size = 10),
    legend.text = ggtext::element_textbox(size = 8),
    legend.key.size = ggplot2::unit(0.4, "cm")
  ) +
  ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size = 0.8)))

plot_env_multi = autoplot(rsmp_env_multi,
  size = 0.3, fold_id = 1, task = task
) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.02)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.02)) +
  ggplot2::theme(
    plot.title = ggtext::element_textbox(
      size = 8,
      r = ggplot2::unit(4, "pt"),
      height = ggplot2::unit(0.30, "inch"), linewidth = 0.5
    ),
    axis.text = ggplot2::element_text(size = 8.5),
    legend.title = ggtext::element_textbox(size = 10),
    legend.text = ggtext::element_textbox(size = 8),
    legend.key.size = ggplot2::unit(0.4, "cm")
  ) +
  ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size = 0.8)))

@@

### Nearest Neighbour Distance Matching --- \code{"spcv_knndm"} {short-title='Nearest Neighbour Distance Matching --- "spcv_knndm"' }

The `spcv_knndm` method from the \pkg{CAST} package uses a two-stage approach: first groups of observations are created using Nearest Neighbour clustering.
These are then merged into k clusters with the aim to minimize the differences between the empirical Nearest Neighbour distribution function of training and prediction points and training and test data during CV using the Wasserstein W statistic [@linnenbrink2023].

The main goal is to ensure similar conditions during CV compared to the actual prediction scenario of the trained model.
The k-fold variant is an extension of the LOOCV variant proposed initially by @mila2022 and more suited for large datasets.

The following example uses the default settings of the implementation in the \pkg{CAST} package.
The method requires a prediction area as an `sf` polygon or point object.
In this case the former was used and passed as an argument to the `"spcv_knndm"` method.

```{r knndm-non-eval, eval=FALSE}
points = sf::st_as_sf(task$coordinates(), crs = task$crs, coords = c("x", "y"))
modeldomain = sf::st_as_sfc(sf::st_bbox(points))

rsmp_knndm = rsmp("spcv_knndm", modeldomain = modeldomain, folds = 5)

autoplot(rsmp_knndm, size = 0.8, fold_id = 1, task = task, label_size = 4)
```

```{r knndm, fig.cap='Random resampling of square spatial blocks using the implementation in package \\pkg{CAST} (method `"spcv\\_knndm"`). Hierarchical clustering has been used together with the "ward.D2" link function and a target of five folds.', out.width="60%", fig.pos="ht", echo=FALSE}
points = sf::st_as_sf(task$coordinates(), crs = task$crs, coords = c("x", "y"))
modeldomain = sf::st_as_sfc(sf::st_bbox(points))

rsmp_knndm = rsmp("spcv_knndm", modeldomain = modeldomain, folds = 5)

suppressMessages(autoplot(rsmp_knndm,
  size = 0.8, fold_id = 1, task = task, label_size = 4
) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.02)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.02)) +
  ggplot2::theme(
    text = ggplot2::element_text(size = 16),
    plot.title = ggtext::element_textbox(
      size = 13,
      r = ggplot2::unit(7, "pt"),
      height = ggplot2::unit(0.46, "inch"), linewidth = 0.8
    )
  ) +
  ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size = 1.5))))
```

@@

```{r, results='hide'}
data = cookfarm_mlr3
set.seed(42)
data$Date = sample(rep(c(
  "2020-01-01", "2020-02-01", "2020-03-01", "2020-04-01",
  "2020-05-01"
), times = 1, each = 35768))
task_spt = as_task_regr_st(data,
  id = "cookfarm", target = "PHIHOX",
  coordinate_names = c("x", "y"), coords_as_features = FALSE,
  crs = 26911
)
task_spt$set_col_roles("Date", roles = "time")

@@

p_lto_print = plotly::layout(p_lto,
  scene = list(camera = list(eye = list(z = 0.58))),
  showlegend = FALSE, title = "",
  margin = list(l = 0, b = 0, r = 0, t = 0)
)

plotly::save_image(p_lto_print, "pdf/lto.pdf",
  scale = 2, width = 1000, height = 800
)
```

@@

p_llo = autoplot(rsmp_cstf_loc,
  fold_id = 5, task = task_spt,
  point_size = 6, axis_label_fontsize = 15,
  plot3D = TRUE, plot_time_var = "Date",
  sample_fold_n = 3000L
)

p_llo_print =
  plotly::layout(p_llo,
    scene = list(camera = list(eye = list(z = 2.5, x = -0.1, y = -0.1))),
    showlegend = FALSE, title = "", polar = TRUE,
    margin = list(l = 0, b = 0, r = 0, t = 0)
  )

plotly::save_image(p_llo_print, "pdf/llo.pdf",
  scale = 2, width = 1000, height = 800
)
```

@@

p_lto = autoplot(rsmp_cstf_time_loc,
  point_size = 6,
  axis_label_fontsize = 15,
  fold_id = 4, task = task_spt, plot3D = TRUE,
  show_omitted = TRUE, sample_fold_n = 3000L
)

p_lto_print = plotly::layout(p_lto,
  scene = list(camera = list(eye = list(z = 0.58))),
  showlegend = FALSE, title = "",
  margin = list(l = 0, b = 0, r = 0, t = 0)
)

plotly::save_image(p_lto_print, "pdf/llto.pdf",
  scale = 2, width = 1000, height = 800
)
```

@@

```{r create-task}
data("ecuador", package = "mlr3spatiotempcv")
task = as_task_classif_st(ecuador,
  target = "slides", positive = "TRUE",
  coordinate_names = c("x", "y"), coords_as_features = FALSE,
  crs = "EPSG:32717"
)
```

@@

```{r vis-spcv-eval, fig.cap="Spatial leave-one-block-out partitioning using coordinate-based clustering to create roughly equally sized polygonal blocks. Due to space limitations only the first two folds of the first repetition are shown.", fig.pos="ht", out.width="100%", echo=FALSE}
autoplot(rsmp_sp, task, fold_id = 1:2, size = 0.3) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.02)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.02)) *
  ggplot2::theme(
    plot.title = ggtext::element_textbox(
      size = 8,
      r = ggplot2::unit(4, "pt"),
      height = ggplot2::unit(0.30, "inch"), linewidth = 0.5
    ),
    axis.text = ggplot2::element_text(size = 8.5),
    legend.title = ggtext::element_textbox(size = 10),
    legend.text = ggtext::element_textbox(size = 8),
    legend.key.size = ggplot2::unit(0.4, "cm")
  ) *
  ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size = 0.8)))
```

@@

For example, buffering, or the use of a spatial or temporal separation distance between training and test sets, is currently only implemented for some methods (`"spcv_buffer"`, `"spcv_disc"`, and `"sptcv_cstf"` with both `space_var` and `time_var`).
Its use should, however, be limited to use cases involving a prediction distance, as a buffer zone reduces the size of the training sample and introduces the risk of geographically biased training data [@meyer2021, @zhu2015].
